The Etiology of Cybersecurity

### 2020-07-19 reports/paper_0/main.tex
Luca

%of sessions such as \autocite{Escobar2007maudenpa}),\fixn{Luca}{Nowadays, most

@@

%\autocite{Oheimb2010aslan++}),\fixn{Luca}{Nowadays, most tools allow for easy

@@

%\autocite{Blanchet2017symbolic}).\fixn{Luca}{This is the only claim that, more

@@

%approach (if any).\fixn{Luca}{I don't think that this is important. In many

@@

\section{Literature Review}\label{sec:literature}
\begin{figure*}[t]
	\centering
	\includegraphics[width=.8\textwidth]{safety-security_3.pdf}
	\caption{Overview of keywords related to security and safety}
	\label{fig:safety-security}
\end{figure*}

In most of the natural languages the concepts of safety
and security are not syntactically differentiated and both terms (safety and
security) are expressed by the same word, e.g. sicurezza in Italian.  A
semantic distinction between safety and security is correlated to a
belief\footnote{A belief has to be intended as a proposition which is supposed
to be true by the majority of people in our society, without a scientific
underlying theory but based on partial empirical evidences or inductive reasoning based on partial empirical evidences.} that
safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
natural environment (e.g. natural events such as wearing of hardware
components) while security deals with \emph{incidents} posed by mankind (e.g.
attackers and bugs).  The fundamental difference between nature and mankind (and,
in turn, between safety and cybersecurity) is believed to be on the different
intents (incidents are intentional while accidents are not) of the causes
that generates a threat; namely, nature is believed not to have malicious
intents (but unfortunate causes-effects) while threats generated by mankind are
malicious\footnote{Of course, logical flaws or bugs may be
introduced by other means (e.g. ignorance) without explicit malicious intents,
but the exploitation of those flaws is considered (for now, and detailed
afterwards in the article) malicious, and then we consider any vulnerability to
be malicious even if due to the lack of skills.}.
This conclusion seems to be against a general formulation of a cybersecurity theory;
how can we define a theory that predicts what a human being will do?
Cybersecurity attacks seems to be related to the creativity of the attacker
and then unpredictable. Currently, our understanding of cybersecurity is
stored into a network of databases of weaknesses (e.g.
CWE\autocite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE\autocite{CVE},
NVD\autocite{NIST2020NVD}),
and attacks (e.g. CAPEC\autocite{MITRE2020CAPEC}, ATT\&CK\autocite{MITRE2020ATTACK}). Systems are
nowadays tested against 
known attacks, ignoring that considering a system secure
if no known attack is possible is against the scientific method.
As an example, if we have a system which is secure against all of the
vulnerabilities in the CVE we just need to wait for a month to have 1000+ more
vulnerabilities to test \autocite{NVD2020dashboard}. If even just one of those new vulnerabilities
affects our system, that falsify the hypothesis that a
system is secure if no known attacks apply. 
We argue that this ``implement and test'' approach should be changed.
Even tough the field 
of security protocols is still lacking of a general cybersecurity theory,
the secure-by-design approach they apply can be considered as going
into the opposite direction of the ``implement and test'' approach.

@@

(therefore should be considered hypothesis and not definitions). While

@@

impossibility of defining a theory that correctly predicts failures.  A failure

@@

the terms that structure our current understanding of security. 

@@

		accord'' (e.g. the physical universe)\autocite{wiki-nature}. 

@@

				equivalent.  From now on, we consider the

@@

		(and adopted in \autocite{nist2013800-53}), is ``weakness in an

@@

		in\autocite{cnssi20104009} may be identified with both Nature

@@

As depicted in Figure~\ref{fig:safety-security}, a vulnerability does not
necessarily become a threat for the system, unless exploited ``through a
channel that allows the violation of the security policy
[\ldots]''\autocite{cnssi20104009}. For example, a software or procedure that takes
advantage of the vulnerability causing an \emph{attack} to the system may
result in several correlated incidents and threats.  The process of
exploitation of a defect as a vulnerability is reported in
Figure~\ref{fig:safety-security} such that the difference between exploit and failure,
and attack and accident, is to be found just in the maliciousness of the intents
that causes this process (i.e. excluding the intent, the terms are just syntactic transformation from a vulnerability to defect, from
accident to incident).

\begin{itemize}
	\item \emph{Weakness}. The definition given by the MITRE
		in \autocite{MITRE2020CWEweakness} of weakness is: ``Software
		weaknesses are errors that can lead to software
		vulnerabilities. A software vulnerability, such as those
		enumerated on the Common Vulnerabilities and Exposures (CVE)
		List, is a mistake in software that can be directly used by a
		hacker to gain access to a system or network''.  The definition
		is circular if we interpret the word ``error'' and ``mistake''
		with the same semantics: a weakness is an error that leads to a
		vulnerability and a vulnerability is a mistake which, in turn,
		is a weakness.  
\end{itemize}

@@

\begin{itemize}
	\item \emph{Causality} refers to the causality principle; defined
		in\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
		connection of phenomena through which one thing (the cause)
		under certain conditions gives rise to, causes something else
		(the effect). The essence of causality is the generation and
		determination of one phenomenon by another. In this respect,
		causality differs from various other kinds of connection, for
		example, the simple temporal sequence of phenomena, of the
		regularities of accompanying processes''. 
		%One can consider the
		%causality principle to be formalized by the K Modal Logic.
	\item An \emph{Exploit}\footnote{We note that the term exploit is only
		used as a verb in\autocite{ISO2009information}} ``[\ldots]
		(from the English verb to exploit, meaning to use something to
		one’s own advantage) is a piece of software, a chunk of data,
		or a sequence of commands that takes advantage of a bug or
		vulnerability to cause unintended or unanticipated behavior to
		occur on computer software, hardware, or something electronic
		(usually computerized).''\autocite{wiki-exploit}.
	\item An \emph{Attack}, as defined by the International Standard
		ISO/IEC 27000 is an ``attempt to destroy, expose, alter,
		disable, steal or gain unauthorized access to or make
		unauthorized use of an asset''; where an \emph{Asset} is
		``anything that has value to the organization''. We note that for
		the purpose of this article, we do not want to focus on a specific
		organization or business to define asset but, in general, on any 
		abstract organization (e.g. a company or a society).
		We do not consider ethical hackers as attacking a system. 
		In fact, we consider the term \emph{hack} as
		non-malicious (as, e.g. in \autocite{Stallman2002hacker}).
	\item A \emph{Threat}, as defined in\autocite{cnssi20104009}, is ``Any
		circumstance or event with the potential to adversely impact
		organizational operations (including mission, functions, image,
		or reputation), organizational assets, individuals, other
		organizations, or the Nation through an information system via
		unauthorized access, destruction, disclosure, modification of
		information, and/or denial of service''.
	\item \emph{Defect}, ``anything that renders the product not reasonably
		safe''\autocite{Robinson2019writing} (i.e. a characteristic of
		an object which hinders its proper usability).
	\item \emph{Failure}, as defined in\autocite{Merriam2020failure} as ``a state of
		inability to perform a normal function''. The term is
		structured and detailed in
		\autocite{cnssi20104009,iet2017glossary} but relying on an
		abstract notion of failure without a specific definition.
	\item \emph{Hazard}, ``a potential source of
		harm''\autocite{iet2017glossary}.
\end{itemize}

@@

Our literature review shows that most of the definitions relates insecurity to
dis-honesty (also called maliciousness or adversarial) of an agent (often
called adversary or attacker). This, however, just moves the problem of
defining what cybersecurity is as the problem of defining what an dishonest
agent can or cannot do in a system\footnote{where an agent is any virtual or
physical entity of the system or using the system (e.g. a device, a software,
or a human being) and dishonesty is not necessarily related to malicious
motivation but also to incompetence or lack of skills.} As in the Dolev-Yao
theory, we may correlate ``being dishonest'' to ``not following the intended
behavior/rules''.  In the case of a generic system, a dishonest agent is,
therefore, any agent that doesn't follow the intended behavior (or
functionality or logic) of the system.  Given the generality of the definition,
and its high level of abstraction, we may conclude that the hypothesis seems
evident. For example, a software can be considered an agent of the system and
whenever it has a bug, it can be exploited causing an Incident. However, this
hypothesis has lead to the un-testable conclusion that the dishonest behaviors
of agents cannot be defined in general (e.g. due to the heterogeneity of agents
and systems) and that huge repository of dishonest behaviors should be kept as
definition, such as CAPEC\autocite{MITRE2020CAPEC}; or that dishonesty of an agent
with respect to a security protocol should be defined as a number of predefined
actions as in
\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
(to name a few). As depicted in Figure~\ref{fig:causality}, in order to define
a theory on cybersecurity and on the malicious behaviors of agents we noticed
that Attack Patterns (or Attacks) are caused-by the presence of Vulnerabilities
in the system. Those Vulnerabilities are, in turn, cause-by the presence of
Weaknesses in the system.  Weaknesses are Errors in the design or
implementation of a system.  Therefore, a theory on cybersecurity should first
predict the Errors in a system design.

>>>>

%of sessions such as \autocite{Escobar2007maudenpa}),\fixnote{Luca}{Nowadays, most

@@

%\autocite{Oheimb2010aslan++}),\fixnote{Luca}{Nowadays, most tools allow for easy

@@

%\autocite{Blanchet2017symbolic}).\fixnote{Luca}{This is the only claim that, more

@@

%approach (if any).\fixnote{Luca}{I don't think that this is important. In many

@@

\section{Literature Review}\label{sec:literature}
\fixnote{Luca}{I think that this section needs thorough rewriting. First of all, you are trying to say too many things, many of which are not really relevant for what you want to say. Second, you are not just reviewing but also introducing some new things that are part of your contributions and this is quite confusing. Third, there are some claims that, in my view, are wrong or at best misleading. Shorten, simplify, aiming to distill really only what is relevant. As it is now, the text is quite confused and confusing.}
\begin{figure*}[t]
	\centering
	\includegraphics[width=.8\textwidth]{safety-security_3.pdf}
	\caption{Overview of keywords related to security and safety}
	\label{fig:safety-security}
\end{figure*}

In most of the natural languages the concepts of safety
and security are not syntactically differentiated and both terms (safety and
security) are expressed by the same word, e.g. sicurezza in Italian.  A
semantic distinction between safety and security is correlated to a
belief\footnote{A belief has to be intended as a proposition which is supposed
to be true by the majority of people in our society, without a scientific
underlying theory but based on partial empirical evidences or inductive reasoning based on partial empirical evidences.} that
safety deals with \emph{accidents} (i.e. an unfortunate incident) posed by the
natural environment (e.g. natural events such as wearing of hardware
components), whereas security deals with \emph{incidents} posed by mankind (e.g.
attackers and bugs). The fundamental difference between nature and mankind (and,
in turn, between safety and cybersecurity\fixnote{Luca}{You spoke of security before, which is much wieder than cybersecurity. This is something you have to be very careful with... political scientists would correctly claim that security is much more than just cyber and that their term is the correct one to contrappose to safety}) is believed to be on the different
intents (incidents are intentional, whereas accidents are not) of the causes
that generates a threat. Namely, nature is believed not to have malicious
intents (but unfortunate causes-effects), whereas threats generated by mankind are
malicious.\footnote{Of course, logical flaws or bugs may be
introduced by other means (e.g. ignorance) without explicit malicious intents,
but the exploitation of those flaws is considered (for now, and detailed
afterwards in the article) malicious, and thus we consider any vulnerability to
be malicious even if it is due to the lack of skills.}.\fixnote{Luca}{Secondo me tutta questa prima discussione \`e alquanto inutile e si pu\`o riassumere in un'unica frase.}
This conclusion seems to be against a general formulation of a cybersecurity theory;
how can we define a theory that predicts what a human being will do?
Cybersecurity attacks seem to be related to the creativity of the attacker
and thus unpredictable. Currently, our understanding of cybersecurity is
stored into a network of databases of weaknesses (e.g.
CWE~\autocite{MITRE2020CWEresearch}), vulnerabilities (e.g. CVE~\autocite{CVE},
NVD~\autocite{NIST2020NVD}),
and attacks (e.g. CAPEC~\autocite{MITRE2020CAPEC}, ATT\&CK~\autocite{MITRE2020ATTACK}).\fixnote{Luca}{Come detto, sono fortemente in disaccordo con questa frase. Parte del nostro understanding \`e l\`i, ma non tutto. Tutto il lavoro teorica e formale non \`e contenuto l\`i. Mi sembra tra l'altro un'affermazione alquanto inutilmente provocatoria} Systems are
nowadays tested against 
known attacks, ignoring that considering a system secure
if no known attack is possible is against the scientific method.\fixnote{Luca}{No, non \`e una scelta ma \`e semplicemente ignoranza di chi fa i test. Chi fa i test seriamente questo lo sa benissimo. E non solo perch\'e \`e contro lo scientific methos ma anche perch\'e testing it not proving}
As an example, if we have a system which is secure against all of the
vulnerabilities in the CVE we often just need to wait for a month to have 1000+ more
vulnerabilities to test \autocite{NVD2020dashboard}. If even just one of those new vulnerabilities
affects our system, that falsifies the hypothesis that a
system is secure if no known attacks apply.\fixnote{Luca}{But, again, the hypothesis was utterly wrong from the start, even without considering the scientific method. No person who is really knowledgeable of testing would ever claim that testing shows the absence of all attacks. Maybe this is what is done in industry, but then industry is simply wrong (it would not be the first time) and people in academia have been saying so for ages. It is not you argueing it.. I believe that building the argument as you do is wrong. I would suggest: (1) Industry uses testing believing that it is enough. (2) It is well know that testing cannot show the absence of attacks. (3) The scientific method makes this even more evident and in fact we show ...} 
We argue that this ``implement and test'' approach should be changed.
Even tough the field 
of security protocols is still lacking of a general cybersecurity theory,
the secure-by-design approach they apply can be considered as going
into the opposite direction of the ``implement and test'' approach.\fixnote{Luca}{Questa ultima frase \`e appesa qui senza una vera chiara connessione a quanto detto finora.}

@@

(therefore should be considered hypotheses and not definitions).\fixnote{Luca}{Hypotheses and definitions don't type check. Moreover, many standards require formal methods proofs, don't they?} While

@@

impossibility of defining a theory that correctly predicts failures.\fixnote{Luca}{This sentence does not follow grammatically}  A failure

@@

the terms that structure our current understanding of security.\fixnote{Luca}{I disagree. It should be cybersecurity, not security. Moreover, as I mentioned, the figure is misleading when it talks about blac-hat hacking. I also don't like ``Manking'', as mankind does many of the things, not just designing systems} 

@@

		accord'' (e.g. the physical universe)~\autocite{wiki-nature}. 

@@

				equivalent.\fixnote{Luca}{Grammar issues}  From now on, we consider the

@@

		(and adopted in \autocite{nist2013800-53}), is a ``weakness in an

@@

		in~\autocite{cnssi20104009} may be identified with both Nature

@@

As depicted in Figure~\ref{fig:safety-security}, a vulnerability does not
necessarily become a threat for the system, unless exploited ``through a
channel that allows the violation of the security policy
[\ldots]''~\autocite{cnssi20104009}. For example, a software or procedure that takes
advantage of the vulnerability causing an \emph{attack} to the system may
result in several correlated incidents and threats.  The process of
exploitation of a defect as a vulnerability is reported in
Figure~\ref{fig:safety-security} such that the difference between exploit and failure,
and attack and accident, is to be found just in the maliciousness of the intents
that causes this process (i.e. excluding the intent, the terms are just syntactic transformation from a vulnerability to defect, from
accident to incident).

\begin{itemize}
	\item \emph{Weakness}. The definition given by the MITRE
		in \autocite{MITRE2020CWEweakness} of weakness is: ``Software
		weaknesses are errors that can lead to software
		vulnerabilities.\fixnote{Luca}{Worth to say that many security issues are not software weakenesses, e.g. cryptography issues or the way cryptography is used. Moreover, in cyber-physical systems, security issues might be due to hardware issues. Again, I think that you are trying to say too many things. Focus on what you really want to say.} A software vulnerability, such as those
		enumerated on the Common Vulnerabilities and Exposures (CVE)
		List, is a mistake in software that can be directly used by a
		hacker to gain access to a system or network''.  The definition
		is circular if we interpret the word ``error'' and ``mistake''
		with the same semantics: a weakness is an error that leads to a
		vulnerability and a vulnerability is a mistake which, in turn,
		is a weakness.  
\end{itemize}

@@

\begin{itemize}
	\item \emph{Causality} refers to the causality principle; defined
		in~\autocite{Spirkin1983Dialectical} as ``Causality is a genetic
		connection of phenomena through which one thing (the cause)
		under certain conditions gives rise to, causes something else
		(the effect). The essence of causality is the generation and
		determination of one phenomenon by another. In this respect,
		causality differs from various other kinds of connection, for
		example, the simple temporal sequence of phenomena, of the
		regularities of accompanying processes''. 
		%One can consider the
		%causality principle to be formalized by the K Modal Logic.
	\item An \emph{Exploit}\footnote{We note that the term exploit is only
		used as a verb in~\autocite{ISO2009information}}\fixnote{Luca}{So what? Is it The use in~\autocite{ISO2009information} is incomplete? } ``[\ldots]
		(from the English verb to exploit, meaning to use something to
		one’s own advantage) is a piece of software, a chunk of data,
		or a sequence of commands that takes advantage of a bug or
		vulnerability to cause unintended or unanticipated behavior to
		occur on computer software, hardware, or something electronic
		(usually computerized).''~\autocite{wiki-exploit}.
	\item An \emph{Attack}, as defined by the International Standard
		ISO/IEC 27000, is an ``attempt to destroy, expose, alter,
		disable, steal or gain unauthorized access to or make
		unauthorized use of an asset''; where an \emph{Asset} is
		``anything that has value to the organization''. We note that for
		the purpose of this article, we do not want to focus on a specific
		organization or business to define asset but, in general, on any 
		abstract organization (e.g. a company or a society).
		We do not consider ethical hackers as attacking a system. 
		In fact, we consider the term \emph{hack} as
		non-malicious (as, e.g. in~\autocite{Stallman2002hacker}).
	\item A \emph{Threat}, as defined in~\autocite{cnssi20104009}, is ``Any
		circumstance or event with the potential to adversely impact
		organizational operations (including mission, functions, image,
		or reputation), organizational assets, individuals, other
		organizations, or the Nation through an information system via
		unauthorized access, destruction, disclosure, modification of
		information, and/or denial of service''.
	\item \emph{Defect}: ``anything that renders the product not reasonably
		safe''~\autocite{Robinson2019writing} (i.e. a characteristic of
		an object which hinders its proper usability).
	\item \emph{Failure}, as defined in\autocite{Merriam2020failure}, is ``a state of
		inability to perform a normal function''. The term is
		structured and detailed in
		\autocite{cnssi20104009,iet2017glossary}, but relying on an
		abstract notion of failure without a specific definition.
	\item \emph{Hazard}, ``a potential source of
		harm''~\autocite{iet2017glossary}.
\end{itemize}

@@

Our literature review shows that most of the definitions relate insecurity to
dis-honesty (also called maliciousness or adversarial) of an agent (often
called adversary or attacker).\fixnote{Luca}{This is slightly misleading, as you have not discussed this above. You mean in the review that you did you discovered that...} This, however, just moves the problem of
defining what cybersecurity is as the problem of defining what a dishonest
agent can or cannot do in a system, where an agent is any virtual or
physical entity of the system or using the system (e.g. a device, a software,
or a human being) and dishonesty is not necessarily related to malicious
motivation but also to incompetence or lack of skills. As in the Dolev-Yao
theory, we may correlate ``being dishonest'' to ``not following the intended
behavior/rules''.  In the case of a generic system, a dishonest agent is,
therefore, any agent that doesn't follow the intended behavior (or
functionality or logic) of the system.  Given the generality of the definition,
and its high level of abstraction, we may conclude that the hypothesis seems
evident.\fixnote{Luca}{Which hypothesis?} For example, a software can be considered an agent of the system and
whenever it has a bug, it can be exploited causing an Incident. However, this
hypothesis has lead to the un-testable conclusion that the dishonest behaviors
of agents cannot be defined in general (e.g. due to the heterogeneity of agents
and systems) and that huge repository of dishonest behaviors should be kept as
definition, such as CAPEC~\autocite{MITRE2020CAPEC}; or that dishonesty of an agent
with respect to a security protocol should be defined as a number of predefined
actions as in
\autocite{Turuani2006clatse,Basin2005ofmc,Armando2016satmc,Rocchetto2017interpolation}
(to name a few).\fixnote{Luca}{I foudn this sentence very hard to follow. Moreover, I would be careful to speak of ``predefined actions'' as basically the DY can do anything so it is not cleaer what you are referring to. Of course, you could this that this excludes, e.g., human errors (say as in the work I did with Diego) or more, but then you should clarify.} As depicted in Figure~\ref{fig:causality}, in order to define
a theory on cybersecurity and on the malicious behaviors of agents we noticed
that Attack Patterns (or Attacks) are caused-by the presence of Vulnerabilities
in the system.\fixnote{``cause-by'' is grammatically wrong, but more importantly the presence of a vulnerability does not necessarily cause an attack, but allows it. Moreover, you speak of Patterns for the first time.} Those Vulnerabilities are, in turn, cause-by the presence of
Weaknesses in the system.  Weaknesses are Errors in the design or
implementation of a system.  Therefore, a theory on cybersecurity should first
predict the Errors in a system design.
